---
title: "Base R code"
format: html
editor: visual
---

## Task 1: Basic Vector practice

**1 & 2** - I created the two vectors, pre_measurements and post_measurements and then I assign names to the vector elements

```{r}
pre_measurements <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)

post_measurements<- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)

names(pre_measurements) <- paste("Subject", 1:20, sep = "_")
names(post_measurements) <- paste("Subject", 1:20, sep = "_")
```

**3** - Here I use the R math element-wise to calculate the blood pressure for each patient

```{r}

change_blood_pressure <- pre_measurements - post_measurements

```

**4** - I am using the mean() to calculate the average decrease in blood pressure across all patients

```{r}

avg_decrease_blood_pressure <- mean(change_blood_pressure)

```

**5** - Using the "which" it will identify which patients experienced a decrease in blood pressure after the treatment and it's a positive change (\>0)

```{r}

patients_decreased <- which(change_blood_pressure > 0)

```

**6** - Indexing so I can filter for positive change

```{r}

positive_change <- change_blood_pressure[change_blood_pressure > 0]

```

**7** - I use the mean to calculate the average of positive change

```{r}

avg_positive_change <- mean(positive_change)

```

## Task 2: Basic Data Frame practice


**1** - Created the dataframe with the four columns (patient, pre_bp, post_bp, diff_bp) 

```{r}

pre_measurements <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
post_measurements <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
change_blood_pressure <- pre_measurements - post_measurements

bp_data <- data.frame(
  patient = paste("Subject", 1:20, sep = "_"),
  pre_bp = pre_measurements,
  post_bp = post_measurements,
  diff_bp = change_blood_pressure
)

```

**2** - Filter the rows where the diff_bp column is negative (<0)

```{r}

negative_diff_bp <- bp_data[bp_data$diff_bp < 0, ]

```

**3** - Added the new column with (True if post_bp < 120) and False if its not

```{r}

bp_data$lower_than_120 <- bp_data$post_bp < 120

```

**4** - Here I just print the data frame

```{r}

knitr::kable(bp_data)

```

## Task 3: List practice

**1** - New information about the treatment that is a placebo, I put those in a new data frame with the additional column and now is 10 subjects rather than 20

```{r}

pre_placebo <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
post_placebo <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
change_placebo <- pre_placebo - post_placebo


placebo_data <- data.frame(
  patient = paste("Subject", 1:10, sep = "_"),
  pre_bp = pre_placebo,
  post_bp = post_placebo,
  diff_bp = change_placebo,
  lower_than_120 = post_placebo < 120
)

```

**2** - The first element of the list is from task 2 and the second element with the new data from task 3

```{r}

blood_pressure_list <- list(
  treatment = bp_data,
  placebo = placebo_data
)

```

**3** - I access the first list element with $ that accesses the name directly, double brackets [[ ]] to access by index and single bracket [ ] to access by name

```{r}

treatment_1 <- blood_pressure_list$treatment

treatment_2 <- blood_pressure_list[[1]]

treatment_3 <- blood_pressure_list["treatment"]

```

**4** - Just accessing pre_bp of the placebo data frame from the previous list 

```{r}

pre_bp_placebo <- blood_pressure_list$placebo$pre_bp

```